Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding

### 2022-05-11 cmsb22.tex
Add a line…
Integrated evaluation into main text

\begin{table}[!htb]
  \caption{Timing comparisons between the proposed method and PyBoolNet on the selected real-world models.
    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively. For each method, Column "\(|M|\)" denotes the number of minimal trap spaces and Column "time (s)" denotes the computation time in seconds.
    "DNF" means that the method did not finish the computation within one hour.}
  \centering
  \label{tab:result_real}
  \begin{tabular}{lrrrrrrr}
    \toprule
    & & & & \multicolumn{2}{c}{PyBoolNet} & \multicolumn{2}{c}{Proposed method}\\
    \cmidrule(lr){5-6}\cmidrule(lr){7-8}
    model & $n$ & $e$ & $s$ & $|M|$ & time (s) & $|M|$ & time (s) \\ \midrule
    inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 287 & 0 & DNF & DNF & 1 & 296.23 \\
    T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 193 & 7 & - & - & 318 & 0.24 \\
    butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 139 & 13 & - & - & 512 & 0.17 \\
    colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 153 & 1 & - & - & 10 & 0.18 \\
    IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 149 & 15 & - & - & 512 & 0.32 \\
    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 246 & 16 & - & - & 5724 & 0.90 \\
    Korkut-2015~\cite{lee2019signal} & 99 & - & 12 & DNF & DNF & 2060 & 2.86 \\ \midrule

    adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 493 & 4 & - & - & 78 & 3.72 \\
    interferon-1~\cite{ostaszewski2021covid19} & 121 & 190 & 55 & - & - & 512 & 0.43 \\
    TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 273 & 2 & - & - & 48 & 0.50 \\
    influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 302 & 11 & - & - & 2560 & 1.07 \\
    signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 431 & 0 & - & - & 812 & 1.40 \\
    HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 368 & 14 & DNF & DNF & 1408 & 1.44 \\
    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 557 & 9 & - & - & 375850 & 331.87 \\
    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 228 & 56 & - & - & 5120 & 0.99 \\
    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 304 & 72 & - & - & 512 & 3.79 \\
    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 268 & 25 & - & - & 1024 & 0.67 \\
    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 506 & 13 & DNF & DNF & 36852 & 14.19 \\
    RA-apoptosis~\cite{aghamiri2020automated} & 181 & - & 59 & DNF & DNF & 640 & 0.70 \\
    er-stress~\cite{ostaszewski2021covid19} & 182 & 266 & 75 & - & - & 25600 & 4.54 \\
    cascade-3~\cite{Tsirvouli2020} & 183 & 603 & 0 & - & - & 1 & 1.65 \\
    CHO-2016~\cite{lee2019signal} & 200 & 696 & 13 & - & - & 1024 & 2.10 \\ \midrule

    T-cell-check-point~\cite{hernandez2020computational} & 218 & - & 34 & - & - & 15360 & 4.33 \\
    ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 1100 & 22 & DNF & DNF & DNF & DNF \\
    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 533 & 19 & - & - & 512 & 1.58 \\
    cholocystokinin~\cite{aghamiri2020automated} & 404 & - & 95 & - & - & 705024 & 1080.35 \\ \midrule

    Alzheimer~\cite{aghamiri2020automated} & 1169 & - & 469 & - & - & 49152 & 66.17 \\
    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & - & 522 & DNF & DNF & 1024 & 1511.07 \\
    human-network~\cite{kim2011reduction} & 1953 & - & 675 & DNF & DNF & 2048 & 1770.98 \\ \midrule

    SN-5~\cite{kim2013rmod} & 2746 & - & 835 & DNF & DNF & 2048 & 3094.04 \\
    turei-2016~\cite{lee2019signal} & 4691 & - & 1257 & DNF & DNF & DNF & DNF \\

    \bottomrule
  \end{tabular}

@@

To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heu=domain --dom-mod=7} used by PyBoolNet). We ran all the benchmarks on a virtual machine whose environment is CPU: Intel® Core™ i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit. Finally, we set a time limit as one hour for each model.

Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.

@@

\appendix

\section{PyBoolNet repo}

%   \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}
%   \vangiang{Indeed, they are not readily available. But if we use more meaningful words, the width of the table may exceed the width of the page. Thus, I use \(e\) and \(s\) along with the explanations shown in the table's caption. Do you have any other option?}
%   \sylvain{just remove them since they are not relevant}

\section{Selected models}

\begin{table}[!htb]
  \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
    Columns \(n\) and \(s\) denote the numbers of nodes (resp.\ input nodes) of each model. Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
    ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of 2 minutes.}
  \centering
  \label{tab:selected_redone}
  \begin{tabular}{lrrrrrr}
    \toprule
    & & & & PyBoolNet & \multicolumn{2}{c}{Trappist} \\
    \cmidrule(lr){6-7}
    model & \(n\) & \(s\) & \(|M|\) & first 1000 & first 1000 & all \\
    \midrule
    inflammatory-bowel-disease & 47 & 0 & 1 & DNF & 21.90 & 25.75\\
    T-LGL-survival-network-2008 & 61 & 7 & 318 & 0.73 & 0.14 & 0.14\\
    butanol-production & 66 & 13 & 8192 & 0.59 & 0.13 & 0.64\\
    colitis-associated-colon-cancer & 70 & 1 & 10 & 0.16 & 0.08 & 0.06\\
    mast-cell-activation & 73 & 19 & 1000+ & 0.64 & 0.15 & DNF\\
    IL-6-signalling & 86 & 15 & 32768 & 3.30 & 0.10 & 5.06\\
    Corral-ThIL-17-diff & 92 & 16 & 1000+ & DNF & 0.38 & DNF\\
    Korkut-2015 & 99 & 12 & 18556 & DNF & 0.61 & 5.01\\
    adhesion-cip-migration-cell-cycle & 121 & 4 & 78 & 40.81 & 0.33 & 0.45\\
    interferon-1 & 121 & 55 & 1000+ & 8.36 & 0.19 & DNF\\
    TCR-TLR5-signaling-2018 & 130 & 5 & 48 & 0.88 & 0.12 & 0.11\\
    influenza-virus-replication-cycle & 131 & 11 & 10128 & 63.62 & 0.21 & 1.95\\
    signaling-in-prostate-cancer & 133 & 11 & 2760 & DNF & 0.32 & 0.72\\
    HIV-1-with-T-cell-signaling & 138 & 14 & 39424 & DNF & 0.87 & 7.72\\
    signal-transduction-in-fibroblasts & 139 & 9 & 1000+ & DNF & 0.56 & DNF\\
    HMOX-1-pathway & 145 & 56 & 1000+ & 5.16 & 0.18 & DNF\\
    kynurenine-pathway & 150 & 72 & 1000+ & DNF & 0.74 & DNF\\
    virus-replication-cycle & 154 & 25 & 1000+ & DNF & 0.19 & DNF\\
    immune-system & 164 & 13 & 1000+ & DNF & 0.67 & DNF\\
    RA-apoptosis & 180 & 59 & 1000+ & DNF & 0.16 & DNF\\
    MAPK & 181 & 37 & 1000+ & 68.62 & 0.64 & DNF\\
    er-stress & 182 & 75 & 1000+ & 28.65 & 0.21 & DNF\\
    cascade-3 & 183 & 0 & 1 & 92.75 &  0.65 & 0.83\\
    CHO-2016 & 200 & 13 & 13312 & DNF & 1.01 & 1.99\\
    T-cell-check-point & 218 & 14 & 1000+ & 37.40 & 0.35 & DNF\\
    ErbB-receptor-signaling & 247 & 22 & 1000+ & DNF & 1.07 & DNF\\
    signaling-macrophage-activation & 321 & 19 & 1000+ & 15.82 & 0.63 & DNF\\
    cholocystokinin & 383 & 74 & 1000+ & 1.22 & 0.73 & DNF\\
    Alzheimer & 762 & 237 & 1000+ & DNF & 1.03 & DNF\\
    KEGG-network & 1659 & 521 & 1000+ & DNF & 44.85 & DNF\\
    human-network & 1953 & 669 & 1000+ & DNF & 42.97 & DNF\\
    SN-5 & 2746 & 829 & ??? & DNF & DNF & DNF\\
    turei-2016 & 4691 & 1257 & ??? & DNF & DNF & DNF\\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet} on both its own repository of models and on large models from the literature.

To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heu=domain --dom-mod=7} used by PyBoolNet). We ran all the benchmarks on an apple laptop whose environment is CPU: Intel\textregistered\ Core\texttrademark\ i7 1.20GHz x 4, 16 GB DDR4 RAM, MacOS 12.3.1. Finally, we set a time limit of two minutes for each model.

\subsection{PyBoolNet repo}

\subsection{Selected models}

We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
Table~\ref{tab:result_real} shows the experimental results on real-world models. Hereafter, we analyze in detail the results with respect to minimal trap space computation.

\begin{table}[!htb]
  \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
    Columns \(n\) and \(s\) denote the numbers of nodes (resp.\ input nodes) of each model. Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
    ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of 2 minutes.}
  \centering
  \label{tab:result_real}
  \begin{tabular}{lrrrrrr}
    \toprule
    & & & & PyBoolNet & \multicolumn{2}{c}{Trappist} \\
    \cmidrule(lr){6-7}
    model & \(n\) & \(s\) & \(|M|\) & 1000 & 1000 & all \\
    \midrule
    inflammatory-bowel-disease~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & 21.90 & 25.75\\
    T-LGL-survival-network-2008~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.73 & 0.14 & 0.14\\
    butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.59 & 0.13 & 0.64\\
    colitis-associated-colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.16 & 0.08 & 0.06\\
    mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & 1000+ & 0.64 & 0.15 & DNF\\
    IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 3.30 & 0.10 & 5.06\\
    Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & 1000+ & DNF & 0.38 & DNF\\
    Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 0.61 & 5.01\\ \midrule
    adhesion-cip-migration-cell-cycle~\cite{guberman2020boolean} & 121 & 4 & 78 & 40.81 & 0.33 & 0.45\\
    interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & 1000+ & 8.36 & 0.19 & DNF\\
    TCR-TLR5-signaling-2018~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 0.88 & 0.12 & 0.11\\
    influenza-virus-replication-cycle~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 63.62 & 0.21 & 1.95\\
    signaling-in-prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & 0.32 & 0.72\\
    HIV-1-with-T-cell-signaling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & 0.87 & 7.72\\
    signal-transduction-in-fibroblasts~\cite{helikar2008emergent} & 139 & 9 & 1000+ & DNF & 0.56 & DNF\\
    HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & 1000+ & 5.16 & 0.18 & DNF\\
    kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & 1000+ & DNF & 0.74 & DNF\\
    virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & 1000+ & DNF & 0.19 & DNF\\
    immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & 1000+ & DNF & 0.67 & DNF\\
    RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & 1000+ & DNF & 0.16 & DNF\\
    MAPK~\cite{aghamiri2020automated} & 181 & 37 & 1000+ & 68.62 & 0.64 & DNF\\
    er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & 1000+ & 28.65 & 0.21 & DNF\\
    cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 92.75 &  0.65 & 0.83\\
    CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 1.01 & 1.99\\
    T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & 1000+ & 37.40 & 0.35 & DNF\\
    ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & 1000+ & DNF & 1.07 & DNF\\
    signaling-macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & 1000+ & 15.82 & 0.63 & DNF\\
    cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & 1000+ & 1.22 & 0.73 & DNF\\
    Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & 1000+ & DNF & 1.03 & DNF\\
    KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & 1000+ & DNF & 44.85 & DNF\\
    human-network~\cite{kim2011reduction} & 1953 & 669 & 1000+ & DNF & 42.97 & DNF\\
    SN-5~\cite{kim2013rmod} & 2746 & 829 & ??? & DNF & DNF & DNF\\
    turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & DNF & DNF\\
    \bottomrule
  \end{tabular}

@@

Note that for all the cases where PyBoolNet did not manage to finish before the timeout, as marked by ``DNF'' in Table~\ref{tab:result_real}, the timeout occurred during the computation of the prime-invariants.
Hence, not even a single minimal trap-space was output by the method.

@@
